---
title: "bone_marrow_assignment"
author: "kritika shrestha"
date: "2024-09-26"
output: html_document
---


```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
library(caret)
library(readr)
library(plotly)
library(GGally)
library(vcd)
library(beeswarm)
```






```{r}
data <- read.csv("bone-marrow.csv")
head(data)
```



#view the structure of the dataset
```{r}
str(data)
```
```{r}
head(data)
```
```{r}
# Checking for missing values
colSums(is.na(data))
```

```{r}
# Impute missing values (usig median For numeric columns)
data$RecipientRh[is.na(data$RecipientRh)] <- median(data$RecipientRh, na.rm = TRUE)
data$CD3dCD34[is.na(data$CD3dCD34)] <- median(data$CD3dCD34, na.rm = TRUE)
data$CD3dkgx10d8[is.na(data$CD3dkgx10d8)] <- median(data$CD3dkgx10d8, na.rm = TRUE)
data$Rbodymass[is.na(data$Rbodymass)] <- median(data$Rbodymass, na.rm = TRUE)
```



```{r}
# For categorical columns, use the mode (most frequent value)
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

data$RecipientABO[is.na(data$RecipientABO)] <- Mode(data$RecipientABO)
data$ABOmatch[is.na(data$ABOmatch)] <- Mode(data$ABOmatch)
data$CMVstatus[is.na(data$CMVstatus)] <- Mode(data$CMVstatus)
data$DonorCMV[is.na(data$DonorCMV)] <- Mode(data$DonorCMV)
data$RecipientCMV[is.na(data$RecipientCMV)] <- Mode(data$RecipientCMV)
data$Antigen[is.na(data$Antigen)] <- Mode(data$Antigen)
data$Alel[is.na(data$Alel)] <- Mode(data$Alel)
data$extcGvHD[is.na(data$extcGvHD)] <- Mode(data$extcGvHD)
```


```{r}
colSums(is.na(data))
```



```{r}
colnames(data)
```

Distribution (Survival Status)
```{r}
# Bar plot for survival status distribution
p <-ggplot(data, aes(x = as.factor(survival_status))) +
  geom_bar(fill = "lightblue", color = "black") +
  ggtitle("Distribution of Survival Status") +
  xlab("Survival Status (0 = No, 1 = Yes)") +
  ylab("Count")
ggplotly(p)




```

 Distribution of Donor Age
```{r}
# Donor Age Histogram
p<-ggplot(data, aes(x = Donorage)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  ggtitle("Distribution of Donor Age") +
  xlab("Donor Age") +
  ylab("Frequency")
ggplotly(p)
```


Distribution of Survival Time
```{r}
# Survival Time Histogram
p <-ggplot(data, aes(x = survival_time)) +
  geom_histogram(binwidth = 100, fill = "salmon", color = "black") +
  ggtitle("Distribution of Survival Time") +
  xlab("Survival Time (days)") +
  ylab("Frequency")
ggplotly(p)
```


Comparing Variables
```{r}
# Box plot to compare survival time by gender match
p<-ggplot(data, aes(x = as.factor(Gendermatch), y = survival_time)) +
  geom_boxplot(fill = c("lightblue", "lightgreen")) +
  ggtitle("Survival Time by Gender Match") +
  xlab("Gender Match (0 = Mismatch, 1 = Match)") +
  ylab("Survival Time (days)")
ggplotly(p)
```


Impact of Stem Cell Source on Survival Time
```{r}
# Box plot for survival time by stem cell source
p<-ggplot(data, aes(x = as.factor(Stemcellsource), y = survival_time)) +
  geom_boxplot(fill = c("orange", "lightblue")) +
  ggtitle("Survival Time by Stem Cell Source") +
  xlab("Stem Cell Source") +
  ylab("Survival Time (days)")
ggplotly(p)

```




Relationships Between Variables
```{r}
# Scatter plot for Donor Age vs. Survival Time
p<-ggplot(data, aes(x = Donorage, y = survival_time)) +
  geom_point(color = "blue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  ggtitle("Donor Age vs. Survival Time") +
  xlab("Donor Age") +
  ylab("Survival Time (days)")
ggplotly(p)
```


Recipient Body Mass vs. Survival Time
```{r}
# Scatter plot for Body Mass vs. Survival Time
p<-ggplot(data, aes(x = Rbodymass, y = survival_time)) +
  geom_point(color = "green", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  ggtitle("Recipient Body Mass vs. Survival Time") +
  xlab("Body Mass (kg)") +
  ylab("Survival Time (days)")
ggplotly(p)
```


Stacked and Grouped Bar Plots
```{r}
# Stacked bar plot for ABO Matching vs. Survival Status
p<-ggplot(data, aes(x = as.factor(ABOmatch), fill = as.factor(survival_status))) +
  geom_bar(position = "fill") +
  ggtitle("ABO Match vs. Survival Status") +
  xlab("ABO Match (0 = No, 1 = Yes)") +
  ylab("Proportion of Survival Status") +
  scale_fill_manual(values = c("red", "blue"), name = "Survival Status")
ggplotly(p)
```



Heatmap for Correlations Between Numeric Variables
```{r}
# Load library for heatmap
library(corrplot)

# Correlation matrix for numeric columns
numeric_cols <- data %>% select(Donorage, survival_time, Rbodymass, CD34kgx10d6, CD3dCD34, CD3dkgx10d8)

# Create correlation matrix
corr_matrix <- cor(numeric_cols, use = "complete.obs")

# Plot the heatmap
corrplot(corr_matrix, method = "color", addCoef.col = "black", tl.col = "black", tl.cex = 0.8)
```


Time to aGvHD (Acute Graft-versus-Host Disease) and Survival Time
```{r}
# Scatter plot for time to aGvHD vs. Survival Time
p<-ggplot(data, aes(x = time_to_aGvHD_III_IV, y = survival_time)) +
  geom_point(color = "purple", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  ggtitle("Time to aGvHD vs. Survival Time") +
  xlab("Time to aGvHD (days)") +
  ylab("Survival Time (days)")
ggplotly(p)
```


Survival Status vs. Donor Age
```{r}
# Box plot for Donor Age vs Survival Status
p<-ggplot(data, aes(x = as.factor(survival_status), y = Donorage)) +
  geom_boxplot(fill = c("lightpink", "lightgreen")) +
  ggtitle("Donor Age vs. Survival Status") +
  xlab("Survival Status (0 = No, 1 = Yes)") +
  ylab("Donor Age")
ggplotly(p)
```


 Survival Status vs. Gender Match
```{r}
# Stacked bar plot for Gender Match vs Survival Status
p<-ggplot(data, aes(x = as.factor(Gendermatch), fill = as.factor(survival_status))) +
  geom_bar(position = "fill") +
  ggtitle("Gender Match vs. Survival Status") +
  xlab("Gender Match (0 = Mismatch, 1 = Match)") +
  ylab("Proportion of Survival Status") +
  scale_fill_manual(values = c("red", "blue"), name = "Survival Status")
ggplotly(p)
```
 
 
Survival Status vs. Recipient Body Mass
```{r}
# Box plot for Body Mass vs Survival Status
p<-ggplot(data, aes(x = as.factor(survival_status), y = Rbodymass)) +
  geom_boxplot(fill = c("lightblue", "lightgreen")) +
  ggtitle("Body Mass vs. Survival Status") +
  xlab("Survival Status (0 = No, 1 = Yes)") +
  ylab("Recipient Body Mass (kg)")
ggplotly(p)
```

 
 Exploring the Effect of CD34 Count on Survival Time
```{r}
# Scatter plot for CD34kgx10d6 vs Survival Time
p<-ggplot(data, aes(x = CD34kgx10d6, y = survival_time)) +
  geom_point(color = "purple", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  ggtitle("CD34 Count vs. Survival Time") +
  xlab("CD34 Count (kgx10d6)") +
  ylab("Survival Time (days)")
ggplotly(p)
```


Donor ABO and Survival Status
```{r}
# Stacked bar plot for Donor ABO vs Survival Status
p<-ggplot(data, aes(x = as.factor(DonorABO), fill = as.factor(survival_status))) +
  geom_bar(position = "fill") +
  ggtitle("Donor ABO vs. Survival Status") +
  xlab("Donor ABO Blood Type") +
  ylab("Proportion of Survival Status") +
  scale_fill_manual(values = c("red", "blue"), name = "Survival Status")
ggplotly(p)
```



Donor-Recipient Age Difference and Survival Status
```{r}
# Calculate Donor-Recipient age difference
data$age_diff <- abs(data$Donorage - data$Recipientage)

# Scatter plot for Age Difference vs Survival Time
p<-ggplot(data, aes(x = age_diff, y = survival_time, color = as.factor(survival_status))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ggtitle("Age Difference vs Survival Time") +
  xlab("Age Difference (Donor - Recipient)") +
  ylab("Survival Time (days)")
ggplotly(p)
```


Exploring Disease Group Impact on Survival
```{r}
# Violin plot for survival time by Disease group
p<-ggplot(data, aes(x = as.factor(Diseasegroup), y = survival_time)) +
  geom_violin(trim = FALSE, fill = "lightblue") +
  geom_boxplot(width = 0.1, fill = "white") +  # Adds a mini box plot inside the violin
  ggtitle("Survival Time by Disease Group") +
  xlab("Disease Group") +
  ylab("Survival Time (days)")
ggplotly(p)
```


Survival Status by GvHD (Graft vs Host Disease) Incidence
```{r}
# Stacked bar plot for GvHD vs Survival Status
p<-ggplot(data, aes(x = as.factor(aGvHDIIIIV), fill = as.factor(survival_status))) +
  geom_bar(position = "fill") +
  ggtitle("GvHD (Grade III-IV) vs. Survival Status") +
  xlab("GvHD Incidence (0 = No, 1 = Yes)") +
  ylab("Proportion of Survival Status") +
  scale_fill_manual(values = c("red", "blue"), name = "Survival Status")
ggplotly(p)
```


```{r}
# Stacked bar plot for Donor CMV vs Recipient CMV and Survival Status
p<-ggplot(data, aes(x = as.factor(DonorCMV), fill = as.factor(RecipientCMV))) +
  geom_bar(position = "fill") +
  ggtitle("Donor CMV vs Recipient CMV and Survival Status") +
  xlab("Donor CMV Status") +
  ylab("Proportion of Survival Status") +
  scale_fill_manual(values = c("red", "blue"), name = "Recipient CMV Status")
ggplotly(p)
```


Survival Status vs. CMV and HLA Mismatches    
```{r}
# Stacked bar plot for CMV Status vs HLA Mismatch and Survival Status
p<-ggplot(data, aes(x = as.factor(CMVstatus), fill = as.factor(HLAmismatch))) +
  geom_bar(position = "fill") +
  ggtitle("CMV Status vs HLA Mismatch and Survival Status") +
  xlab("CMV Status") +
  ylab("Proportion of Survival Status") +
  scale_fill_manual(values = c("red", "blue"), name = "HLA Mismatch Count")
ggplotly(p)
```



Survival Time Distribution by Multiple Factors
```{r}
# Faceted plot for survival time by multiple factors
p<-ggplot(data, aes(x = survival_time)) +
  geom_histogram(binwidth = 100, fill = "steelblue", color = "black") +
  facet_grid(Gendermatch ~ ABOmatch) +
  ggtitle("Survival Time by Gender Match and ABO Match") +
  xlab("Survival Time (days)") +
  ylab("Count")
ggplotly(p)
```





Density Plot of Survival Time by Categorical Variables
```{r}
# Density plot for Survival Time by Gendermatch
p <-ggplot(data, aes(x = survival_time, fill = as.factor(Gendermatch))) +
  geom_density(alpha = 0.5) +
  ggtitle("Density Plot of Survival Time by Gender Match") +
  xlab("Survival Time (days)") +
  ylab("Density") +
  scale_fill_manual(values = c("red", "blue"), name = "Gender Match")
ggplotly(p)
```


Bubble Plot
```{r}
# Bubble plot for Donor Age, Rbodymass, and CD34 count with bubble size as survival time
p<-ggplot(data, aes(x = Donorage, y = Rbodymass, size = survival_time, color = as.factor(survival_status))) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(range = c(1, 10)) +
  ggtitle("Bubble Plot: Donor Age, Body Mass, and Survival Time") +
  xlab("Donor Age") +
  ylab("Recipient Body Mass")
ggplotly(p)
```


